---
title: "Assignment 5"
author: "Anil Kemisetti"
date: "9/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
For this assignment you will perform EDA (Exploratory Data Analysis) and create a linear
regression model of data that has been simulated to relate baby IQ to family
characteristics. The dataset may be found in file baby_IQ_data200.csv in the assignment 5
folder (also linked from Assign 5 on Canvas). Variables are described below.

####Parent Characteristics 

  - MA: Maternal Age at time of pregnancy (years) -- maternal_age
  - PSS: Social Status of family -- social_status
  - BE: Breadwinners Post-high school Education -- education
  - CC: cigarette consumption mother per day -- noof_cigarette
  - NP: No. of pregnancies of mother, including this baby -- noof_pregnancies
   
####Infant Characteristics
   - GA: Estimated gestational age of baby at birth (weeks) -- gestational_age
   - BW: Birth weight (g) -- birth_weight
   - BL: Birth height (cm) -- birth_height
   - IQ: IQ score of the baby -- IQ
  
```{r message=FALSE, warning=FALSE}
setwd("~/Documents/usfca/631")
library(ggplot2)
library(dplyr)
library(kableExtra)
library(data.table)
library(psych)
library(tidyr)
library(car)
library(purrr)
library(ggpubr)
library(gridExtra)
library(grid)
library(corrplot)
library(GGally)
library(ggfortify)
```
Reading the datafile.
```{r}
babies <- read.csv("baby_IQ_data200.csv")
head(babies) %>% 
  kable(escape = F, align = "c") %>%
  kable_styling(c("striped","condensed"), full_width = T)
```
Renaming the columnNames
```{r}
setnames(babies, c("MA","PSS","BE","CC","NP","GA","BW","BL","IQ"),c("maternal_age","social_status","education","noof_cigarette","noof_pregnancies",
                                                                    "gestational_age","birth_weight","birth_height","IQ"))
```

```{r}
describe(babies)  %>% 
  kable(escape = F, align = "c") %>%
  kable_styling(c("striped"), full_width = T)
```
function to display plots in a grid
```{r}
grid9x9 <- function(plist,cols, rows, n ){
  layout <- matrix(seq(1,n),ncol = cols, nrow=rows)
  grid.newpage()
  pushViewport(viewport(layout = grid.layout(rows,cols)))
  for(i in 1:9){
    idx <- as.data.frame(which(layout ==i, arr.ind = TRUE))
    print(plist[[i]], vp = viewport(layout.pos.row = idx$row,layout.pos.col = idx$col))
  }
}
```
####Univariate Plot
Following are the various density plots and their mean for all of the variables. 

```{r fig.height=10, fig.width=10, message=FALSE}
p <- list()
coln <- colnames(babies)
for (i in 1:ncol(babies)) {
  p[[i]] <- ggdensity(babies, x=coln[i], fill="#0073C2FF", rug = TRUE, add="mean")
}
grid9x9(p,3,3,9)
```
####BiVariate Plots for all the variables
```{r}
cor(babies)  %>% 
  kable(escape = F, align = "c") %>%
  kable_styling(c("striped"), full_width = T)
```

```{r fig.height=15, fig.width=15}
corrplot(cor(babies), method="pie", title="Correlation plot", mar=c(0,0,5,0))
```
Above plot shows that IQ is correlated postitively with maternal_age and social_status and negatively correlated with noof_cigarattes.

Let us do a pair plot and understand this further. 
```{r fig.height=15, fig.width=15, message=FALSE}
ggpairs(babies, diag = list(continious = "barDiag")) + ggtitle("Bi Variate plot of all the variables")
```
When it comes to Maternal Age and IQ. Maternal age is beteen 25 and 35 and it showed a postive correlation. Similar is the case with social status. 
Majority of values fall inside the median. In the case of Cigaratte Somking. Extreme smoking caused low IQ. 

Let us now explore multiple regression with IQ and all other variables as main effect

```{r fig.height=15, fig.width=15}
fit <- lm(IQ ~ maternal_age+social_status+noof_cigarette+education+noof_pregnancies+gestational_age+birth_weight+birth_height, data = babies)
autoplot(fit)
```
Above plots show a parabola pattern for fitted and residual which is an indication of nonlinear fit. 
Getting the summary details
```{r}
summary(fit)
```
Above sumarry showed that maternal age, noof_cigaratte, education as important variables. 
Using Step function

```{r}
step(lm(IQ ~ maternal_age+social_status+noof_cigarette+education+noof_pregnancies+gestational_age+birth_weight+birth_height, data = babies),direction="both")
```
Step function has figured out that four variable maternal_age, social_status, noof_cigarette and education are important. 
let us fit a model and see the fitted vs residaul plot for these

```{r fig.height=15, fig.width=15}
fit <- lm(IQ ~ maternal_age+social_status+noof_cigarette+education, data = babies)
autoplot(fit)
```
```{r}
summary(fit)
```
In this case social status is not significant. May be it just had interaction effect. 
Fitting main effects and 2nd order interactions of these 3 variables and including ineraction for social status
```{r fig.height=15, fig.width=15}
fit1 <- lm(IQ ~ (maternal_age+noof_cigarette+education)^2 + I(maternal_age^2)+I(noof_cigarette^2)+I(education^2), data=babies)
autoplot(fit1)
```

```{r}
summary(fit1)
```
All the main effects and 2nd order interactions are significant. Got a good adjusted R squared. A nice fit. Now let us explore the ineractions of social staus

```{r fig.height=15, fig.width=15}
fit1 <- lm(IQ ~ (maternal_age+noof_cigarette+education)^2 + I(maternal_age^2)+I(noof_cigarette^2)+I(education^2)+ education:social_status, data=babies)
autoplot(fit1)
```
```{r}
summary(fit1)
```
A slight improvement in the Adjusted r value but this model showed that education and social status interaction is important. 
Let us explore its interaction with maternal age staus. 

```{r fig.height=15, fig.width=15}
fit1 <- lm(IQ ~ (maternal_age+noof_cigarette+education)^2 + I(maternal_age^2)+I(noof_cigarette^2)+I(education^2)+ maternal_age:social_status, data=babies)
autoplot(fit1)
```
```{r}
summary(fit1)
```
This interaction is also important
Let us include all of three main variables with social status

```{r fig.height=15, fig.width=15}
fit1 <- lm(IQ ~ (maternal_age+noof_cigarette+education)^2 + I(maternal_age^2)+I(noof_cigarette^2)+I(education^2)+ maternal_age:social_status+education:social_status+noof_cigarette:social_status, data=babies)
autoplot(fit1)
```
```{r}
summary(fit1)
```
When all of them are included it only selected noof_cigarette:social_status. So the winning model would be
the following

```{r fig.height=15, fig.width=15}
fit1 <- lm(IQ ~ (maternal_age+noof_cigarette+education)^2 + I(maternal_age^2)+I(noof_cigarette^2)+I(education^2)+ noof_cigarette:social_status, data=babies)
autoplot(fit1)
```
```{r}
summary(fit1)
```
## Winning Model
 IQ ~ (maternal_age + noof_cigarette + education)^2 +  I(maternal_age^2) + I(noof_cigarette^2) + I(education^2) + noof_cigarette:social_status
 with an ajusted R-Squared value of 0.9389 
 
