---
title: "Assignment 6"
author: "Anil Kemisetti"
date: "10/5/2018"
output: 
  html_document :
     toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Assignment Description
For this assignment you will investigate the data in file weather.data.csv in the assignment
6 folder (also linked from Assign 6 on Canvas). You will fit a liner model to the raw data
to model the relationship between high daily temperature and month. Then you will create
a copy of the raw data and perform data preparation on it. You will then fit a linear model
to model the relationship between high daily temperature and month in the new dataset and
compare the two models.
In the plots you produce, be sure to include plot title and axis labels (including units where
appropriate).

#Initial Setup 
##Loading the Libraries
```{r message=FALSE, warning=FALSE}
setwd("~/Documents/usfca/631")
library(ggplot2)
library(dplyr)
library(kableExtra)
library(data.table)
library(psych)
library(tidyr)
library(car)
library(purrr)
library(ggpubr)
library(gridExtra)
library(grid)
library(corrplot)
library(GGally)
library(ggfortify)
library(cowplot)
library(ggstatsplot)
library(nortest)
library(ggfortify)
library(ggcorrplot)
library(broom)
library(jtools)
library(RColorBrewer)
library(paletteer)
```
#Exploratory Data Analysis

***
>###Assignment Section
>1. Read the file weather

***

## Reading the data and looking at it for the first time.
### Numerical EDA


```{r fig.width=15, fig.height=5}
temper <- read.csv("weather.data.csv")
p1 <- ggtexttable(summary(temper))
p2 <- ggtexttable(describe(temper))

annotate_figure(plot_grid(p1,p2, nrow=2) ,
               top = text_grob("Summarizing the temper dataframe \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 1"
)

```

***

>### Assignment Section
>2. Perform EDA: Examine the environment and execute a summary of the dataset.
>3. Perform EDA: Create plots of the distribution of each variable.
>4. Perform EDA: Execute the command plot(temper) to plot the pairwise relations of
>the variables in temper. Note that you do not need to add title and axis labels to the
>plot produced by this command.
>5. In order to produce a linear model of upper temperature as a function of month,
>some data preparation will be required. Consider whether the data types of the
>variables appropriately represent temperature, rainfall, month and year. Write a
>brief comment identifying any variable that is typed incorrectly. Do not modify
>dataset temper.
>#### Analysis
>Initial look at the data shows that month and year are integers. They are discrete values. They should be factor variables
>Please find below the EDA for the above set of questions

***

### Visual EDA
#### Univatiate Analysis
Visualing the continious variables

```{r fig.height=15, fig.width=15, message=FALSE, warning=FALSE, paged.print=FALSE}
p1 <- ggplot(temper, aes(x=upper))+ geom_histogram( colour="black", aes(y=..density.., fill=..count..))+
      geom_density(linetype = "dashed", color="yellow", size=2)+
      scale_fill_gradient("Count", low="lightblue", high="darkblue") +stat_function(fun=dnorm,color="red", size=1,
                         args=list(mean=mean(temper$upper), 
                                  sd=sd(temper$upper))) +ggtitle("Histogram and Density plot for count and density \n of Upper Temperature") + theme_gray()

p2 <- ggplot(temper, aes(x=lower))+ geom_histogram( colour="black", aes(y=..density.., fill=..count..))+
      geom_density(linetype = "dashed", color="yellow", size=2)+
      scale_fill_gradient("Count", low="lightblue", high="darkblue") +stat_function(fun=dnorm,color="red", size=1,
                         args=list(mean=mean(temper$lower), 
                                  sd=sd(temper$lower))) +ggtitle("Histogram and Density plot for count and density \nof Lower Temperature") + theme_gray()

p3 <- ggplot(temper, aes(x=rain))+ geom_histogram( colour="black", aes(y=..density.., fill=..count..))+
      geom_density(linetype = "dashed", color="yellow", size=2)+
      scale_fill_gradient("Count", low="lightblue", high="darkblue") +stat_function(fun=dnorm,color="red", size=1,
                         args=list(mean=mean(temper$rain), 
                                  sd=sd(temper$rain)))  +ggtitle("Histogram and Density plot for count and density \nof Rain") + theme_gray()

p4 <-gghistostats(
     data = temper,   
     x = upper,                                  
     title = "Distribution of Upper Temperature",   
    fill.gradient = TRUE,                     
    messages = FALSE,
    ggstatsplot.layer = FALSE
) +theme_gray()

p5 <-gghistostats(
     data = temper,   
     x = upper,                                  
     title = "Distribution of Lower Temperature",   
    fill.gradient = TRUE,                     
    messages = FALSE,
    ggstatsplot.layer = FALSE
) +theme_gray()

p6 <-gghistostats(
     data = temper,   
     x = upper,                                  
     title = "Distribution of Rain",   
    fill.gradient = TRUE,                     
    messages = FALSE,
    ggstatsplot.layer = FALSE
) +theme_gray()

p7 <- ggplot(temper, aes(sample = temper$upper))+stat_qq() + stat_qq_line(color = "red") +ggtitle("Testing the  normality for Upper Temperature") +theme_gray()
p8 <- ggplot(temper, aes(sample = temper$lower))+stat_qq() + stat_qq_line(color="red") +ggtitle("Testing the  normality for Lower Temperature") +theme_gray()
p9 <- ggplot(temper, aes(sample = temper$rain))+stat_qq() + stat_qq_line(color ="red") +ggtitle("Testing the  normality for Rain") +theme_gray()

annotate_figure(plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9, ncol=3) ,
               top = text_grob("Visualing Univatiate plot for Continious Variable \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 2"
)
```
Normality test for the continious Variables

```{r fig.width=8, fig.height=3}
ad_test_df <- function(df, alpha= 0.05) {
  l <- lapply(df, ad.test)
  s <- do.call("c", lapply(l, "[[", 1))
  p <- do.call("c", lapply(l, "[[", 2))
  sig <- ifelse(p > alpha, "Passed Normality", "Failed Normality")

  return(list(statistic= s,
              p.value= p,
              significance= sig
              ))
}
t <- ad_test_df(temper)
p1 <- ggtexttable(do.call(rbind,t))
annotate_figure(plot_grid(p1, ncol=1) ,
               top = text_grob("Normality test for the continious variables \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 3")
```
Numerical Description of the discrete variables

```{r fig.height=4, fig.width=12}
message_function <- function(){'Variable is not discrete ignoring'}
cat_data <- function(x){
  frequency_count <- table(x)
  proportion <- round(frequency_count/sum(frequency_count),3)
  percent <- round(proportion*100,3)
  return (rbind(frequency_count,proportion,percent))
}
t <- lapply(temper, function(x) {
    if (is.integer(x)){ return(cat_data(x))}
    else{ return(message_function())}
})
p1<- ggtexttable(t$month)
p2 <- ggtexttable(t$yr)

annotate_figure(plot_grid(p1,p2, ncol=1) ,
               top = text_grob("Proportions for the discrete variables \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 4"
)
```
Visualizing the Discrete variables
```{r fig.height=15, fig.width=10}
p1 <- ggplot(temper, aes(month)) + geom_bar() + ggtitle("Frequency counts for month") +theme_gray()
p2 <- ggplot(temper, aes(yr)) + geom_bar() + ggtitle("Frequency counts for years")+theme_gray()
p3 <- ggpiestats( data = temper,
  main = month,
  title = "Composition of months",
  messages = FALSE
)+theme_gray()
p4 <- ggpiestats( data = temper,
  main = yr,
  title = "Composition of year",
  messages = FALSE
)+theme_gray()
p5 <- ggplot(temper, aes(x=month)) + geom_bar(aes(y=(..count..)/sum(..count..)))+ scale_y_continuous(labels=scales::percent) +ylab("Percentages") + ggtitle("Percentages for month") +theme_gray()
p6 <- ggplot(temper, aes(x=yr)) + geom_bar(aes(y=(..count..)/sum(..count..)))+ scale_y_continuous(labels=scales::percent) +ylab("Percentages") + ggtitle("Percentages for month")+theme_gray()

annotate_figure(plot_grid(p1,p2,p3,p4,p5,p6, ncol=2) ,
               top = text_grob("Proportions for the discrete variables \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 5"
)
```

#### Bivariarte Analysis
Correlation Analysis of the bi-variate analysis.
Correlation Matrix of all the variables

```{r fig.height=3, fig.width=10}
p1 <- ggtexttable(cor(temper)) + ggtitle("Correlation matrix") 

annotate_figure(plot_grid(p1, ncol=1) ,
               top = text_grob("Correlation matrix of all numeric variables \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 6"
)

```
```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
p1 <- ggcorrplot(cor(temper), hc.order = TRUE,  type = "lower",
       outline.col = "white",
       method = "circle",
       lab = TRUE,
       ggtheme = ggplot2::theme_gray,
       colors = c("#6D9EC1", "white", "#E46726")) + ggtitle("Correlalogram for months")
capture.output(p2 <- ggcorrmat(
  data =temper,
  corr.method = "robust",                   
  sig.level = 0.001,                         
  p.adjust.method = "holm",                  
  cor.vars = c(upper,lower,rain,month,yr),    
  matrix.type = "upper",                     
  colors = c("#6D9EC1", "white", "#E46726"), 
  title = "Correlalogram for months",
  messages = FALSE
) +theme_gray(),file = 'NUL')

annotate_figure(plot_grid(p1,p2, nrow = 1) ,
               top = text_grob("Correlalogram of all numeric variables \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 7"
) 

```
Plotting the regression for every pair of the variables. 

```{r fig.height=30, fig.width=30, message=FALSE, warning=FALSE}
regression_fn <- function(data, mapping, method, ...){
      ggplot(data = data, mapping = mapping) + 
      geom_point() + 
      geom_smooth(method=method, ...)
    }
p1 <- ggpairs(temper, diag = list(continious = "barDiag"), lower= list(continuous = wrap(regression_fn, method="lm") ), upper= list(continuous = wrap(regression_fn, method = "loess") )) + ggtitle("Bi Variate plot of all the variables") + theme_gray()
f <- ggmatrix_gtable(p1,nrow=1)
capture.output(annotate_figure(f ,
               top = text_grob("Regressiion plot of all pairs of variables \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 8"
),file = 'NUL')
```
***

>### Assignment Section
>6. Use the lm command to produce a linear regression model of upper temperature as a
>function of month for temper. Store your model in a variable named fit0.
>7. Execute a summary of fit0 and briefly describe its goodness of fit. Be sure to
>include what quantity in the summary leads to your statement.
>8. Execute plot(fit0,1). Note that a parabolic shape of the plot Residuals vs Fitted
>often, but not always, is indicative of 2nd order effects.
>9. Write out the equation of the fitted line for model fit0.
>10. Write out what the regression line of fit0 models: what is the effect on response
>variable upper of a unit increase in predictor variable month?
>11. Create a scatter plot of upper temperature as a function of month in dataset
>temper. Color the points by month.
>12. On your plot add an abline of the fitted regression line whose equation you wrote
>above. You may use the command abline(lm(fit0)) to accomplish this.

>Please find the answers below

***

#Regression Analysis
Performing regression analysis of the model fitting the upper as a function of month and creating the summary of the model.

```{r fig.height=7, fig.width=17}
fit0 <- lm(upper~month, data = temper)
t <- tidy(fit0, conf.int = TRUE)
p1 <- ggtexttable(t)
p2 <- ggtexttable(head(augment(fit0)))
p3 <- ggtexttable(glance(fit0))

annotate_figure(plot_grid(p1,p2,p3, nrow=3) ,
               top = text_grob("Summarizing the fit of the model upper~month \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 9"
)

```

Plotting the model details
```{r fig.height=5, fig.width=15}
p1 <- ggplot(temper, aes(x=month,y = upper, color=month))+ geom_point() + geom_abline(slope = t$estimate[2], intercept = t$estimate[1],color="red",size=1)  + ggtitle("Plotting Regression Line") + theme_gray()
p2 <- ggplot(t, aes(estimate, term, color = term)) + geom_point() + geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +geom_vline(xintercept = 0)  + ggtitle("Plotting Regression coefficients") + theme_gray()
p3 <- ggcoefstats(x = fit0) + ggtitle("Plotting Regression coefficients") + theme_gray()

annotate_figure(plot_grid(p1,p2,p3, nrow=1) ,
               top = text_grob("Plotting the regression model \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 10"
)
```
Plotting the results of Regression

```{r fig.height=10, fig.width=15}
p <- autoplot(fit0, ncol =4) + theme_gray()
p1 <- plot_summs(fit0, scale = TRUE, plot.distributions = TRUE, inner_ci_level = .9) + theme_gray() +ggtitle("Plot coefficient uncertainty as normal distributions")

annotate_figure(plot_grid(p[[1]],p[[2]],p[[3]],p[[4]],p1,ncol = 3),
               top = text_grob("Plotting the regression results \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 11"
)
```
```{r}
cat('Equation for fit0 is upper = ',t$estimate[2],'* month + ',t$estimate[1] )
```
***

### Assignment Section
>13. Create a new dataset temper2 that is a copy of temper. Then modify temper2 so
>as to correct the data types of variable(s) that are of inappropriate type. This will
>prepare the data for fitting a linear model of upper temperature as a function of
>month. If you are converting a variable to a categorical, create an unordered factor
>variable, as the resulting linear model will be more interpretable. However, be sure
>to specify the output order of the levels in a logical order using “levels=”
>14. Create a plot to visualize upper temperature by month in dataset temper2. Use a
>plot type that is appropriate for the types of these 2 variables in temper2. Briefly
>compare this plot to the one you created above to visualize upper temperature as a
>function of month in dataset temper2.
>15. Use the lm command to produce a linear regression model of upper temperature as a
>function of month for temper2. Store your model in a variable named fit1.
>16. Execute plot(fit1,1). Compare this plot to the Residuals vs Fitted plot of fit0.
>17. Execute a summary of fit1 and briefly describe its goodness of fit in comparison
>with fit0.
>18. Write out three equations of the fitted line for model fit1: one for the month of
>January, one for June, and one for December.

***

# Modifying the dataset
Initial observation was made that month and year are not discrete numbers but are factor variables. Converting them into factor variables.
```{r}
temper2 <- temper
lab = c("Jan", "Feb", "Mar","Apr", "May", "Jun","July", "Aug", "Sep","Oct", "Nov", "Dec")
temper2$month <- factor(temper$month, labels=lab)
temper2$yr <- factor(temper$yr)
```

Describing temper2 numerically
```{r fig.width=15, fig.height=5}
p1 <- ggtexttable(summary(temper2))
p2 <- ggtexttable(describe(temper2))

annotate_figure(plot_grid(p1,p2, nrow=2) ,
               top = text_grob("Summarizing the temper dataframe \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 11"
)
```

Plotting the regression for every pair of the variables. 

```{r fig.height=30, fig.width=30, message=FALSE, warning=FALSE}

p1 <- ggpairs(temper2, diag = list(continious = "barDiag"), lower= list(continuous = wrap(regression_fn, method="lm") ), upper= list(continuous = wrap(regression_fn, method = "loess") ), cardinality_threshold = 19) + 
  ggtitle("Bi Variate plot of all the variables") + theme_gray()
f <- ggmatrix_gtable(p1,nrow=1)
capture.output(annotate_figure(f ,
               top = text_grob("Regressiion plot of all pairs of variables \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 12"
),file = 'NUL')
```

#Regression Analysis
Performing regression analysis of the model fitting the upper as a function of month and creating the summary of the model.

```{r fig.height=10, fig.width=17}
fit1 <- lm(upper~month, data = temper2)
t1 <- tidy(fit1, conf.int = TRUE)
p1 <- ggtexttable(t1)
p2 <- ggtexttable(head(augment(fit0)))
p3 <- ggtexttable(glance(fit0))

annotate_figure(plot_grid(p1,p2,p3, nrow=3) ,
               top = text_grob("Summarizing the fit of the model upper~month \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 13"
)

```


Plotting the model details
```{r fig.height=5, fig.width=15}
color <- grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]
slp <- c(0,t1$estimate[2:12])
slp
mon <- c(1,2,3,4,5,6,7,8,9,10,11,12)
mon
ccc <- sample(color, 12)
incpt=rep(t1$estimate[1],12)
slp*mon
incpt_calc <- (slp*mon)+incpt
color_dt <- data.frame(month=lab,ccc=ccc,slp=slp,incpt=incpt,mon=mon,incpt_calc=incpt_calc)
color_dt
temper3 <- merge(temper2,color_dt,by="month")
p1 <- ggplot(temper3, aes(x=month,y = upper,color=month))+ geom_point() + geom_abline( mapping = aes(slope = 0, intercept = incpt_calc,color=month),size=1)  + ggtitle("Plotting Regression Line") + theme_gray()
p2 <- ggplot(t1, aes(estimate, term, color = term)) + geom_point() + geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +geom_vline(xintercept = 0)  + ggtitle("Plotting Regression coefficients") + theme_gray() 

annotate_figure(plot_grid(p1,p2, nrow=1) ,
               top = text_grob("Plotting the regression model \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 14"
)
```
Plotting the results of Regression

```{r fig.height=10, fig.width=15}
p <- autoplot(fit1, ncol =4) + theme_gray()
p1 <- plot_summs(fit1, scale = TRUE, plot.distributions = TRUE, inner_ci_level = .9) + theme_gray() +ggtitle("Plot coefficient uncertainty as normal distributions")
p2 <- plot_summs(fit0, fit1, scale = TRUE) +ggtitle("Comparing the fit of the two models")+ theme_gray()

annotate_figure(plot_grid(p[[1]],p[[2]],p[[3]],p[[4]],p1,p2,ncol = 3),
               top = text_grob("Plotting the regression results \n", color="darkred", face = "bold", size = 14),
               bottom = text_grob("Data source: \n weather.data data set", color = "blue",
                                  hjust = 1, x = 1, face = "italic", size = 10),
               fig.lab = "Figure 11"
)
```
creating equations for all the months
```{r}
call_func <- function(x) cat('Equation for month ',x[1],' is ',x[4],'+',x[3],'*',x[5],"\n")
apply(color_dt,1,call_func)
```
### r-Squared of the first model was 0.0307 and it improved in the second model and the adjusted R-squared value is 0.0305. Even though the r-squared has not improved there is not explained variance let in the Residual Vs Fitted Plot. 

